|

### 2021-11-25 03-Chp-Systematic-Review-Methods.Rmd
End of day:

# Systematic review of all available evidence on the association between blood lipids and dementia outcomes {#sys-rev-methods-heading}

@@

<!-- TODO Consider reordering the forest plots so that they are the other way around -->

<!-- TODO Extra studies to include: -->
<!-- * Larsson - 10.1136/bmj.j5375 [@larsson2017] -->

@@

This chapter presents a systematic review of primary studies that have examined the relationship between the levels of blood lipids (such as cholesterol and triglycerides), and treatments that change these levels such as statins, and dementia outcomes.

My review included `r n_included` primary studies that contained information on this relationship. I found that statins appear to reduce the risk of Alzheimer's disease, but had no effect of vascular dementia. Lipid levels were not associated with any outcome. The methods used in some of the primary studies meant that I was less confident in the accuracy of their results.

The added value of including preprinted study reports, made possible using the tool described in the previous chapter, along with the use of the results of this review in subsequent chapters, is discussed.

<!-- TODO Search for MCI -->

@@

In this chapter, I describe a comprehensive systematic review of the relationship between blood lipid levels, and treatments that modify them, and the subsequent risk of dementia and related outcomes.

This analysis sought to address two specific aims. Firstly, as discussed in the Introduction to this thesis (Section \@ref(evidence-association)), several diverse forms of evidence on the relationship of lipids and dementia exist. These include randomised controlled trials, observational studies of different analytical design, and Mendelian randomisation studies. However, based on a scoping review of existing literature, no previous evidence synthesis exercise has attempted to examine the association of lipids/statins with dementia outcomes across these distinct evidence types. Collating these diverse evidence sources is important, as if the observed association between lipids and dementia is constant across them, it increases our confidence in the association. As such, the primary aim of this analysis was to systematically review all available literature describing prospective analyses, regardless of study design.

@@

The results of this review are used to guide the primary analysis presented in Chapter \@ref(cprd-analysis-heading) and \@ref(ipd-heading), in addition to forming a key evidence source used in the triangulation exercise presented in Chapter \@ref(tri-heading).

@@

A pre-specified protocol for this analysis was registered on the Open Science Framework platform and is available for inspection.[@mcguinnessluke2020] Deviations from this protocol are detailed in the relevant sections.

@@

In line with best-practice guidance, secondary reviewers were used to check the accuracy of screening, data extraction and risk-of-bias assessment processes. Due to the scale of the project, this review was performed in conjunction with a team of secondary reviewers (see Acknowledgements and Author Declaration in the front matter). <!-- TODO add the following to acknowledgements These included Alexandra MacAleenan, Athena Sheppard, and Matthew Lee.  In addition, Sarah Dawson, an information specialist, provided input to the design of the search strategy.-->

<!-- TODO Double check this list at the end of the project. -->

@@

### Search strategy

@@

Grey literature was searched via ProQuest, OpenGrey and Web of Science Conference Proceedings Citation Index, while theses were accessed using the Open Access Theses and Dissertations portal. In addition, the abstracts list of relevant conferences (e.g. the proceedings of the Alzheimer's Association International Conference, published in the journal Alzheimer's & Dementia) were searched by hand. <!-- TODO how were these searched? Date and other range --> Finally, the reference lists of included studies were searched by hand while studies citing included studies was examined using Google Scholar (forward and reverse citation searching or "snowballing"<!-- TODO CITATION NEEDED -->).

@@

The criteria against which records were assessed for eligibility are presented in the subsequent sections.

I sought to include studies that examined blood lipid levels as a risk factor for demenita outcomes, defined either as binary hypercholesterolemia variable or by category/1-standard-deviation increase of a specific lipid fraction (total cholesterol, high- and low-density lipoprotein cholesterol and triglycerides). I also aimed to include studies examining the effect of treatments that modify lipids levels as a source of indirect evidence. Eligible study designs included randomized controlled trials and non-randomized observational studies of lipid modifying treatments, longitudinal studies examining the effect of increased/decreased blood lipid levels, and genetic instrumental variable (Mendelian randomization) studies examining the effect of genetically increased/decreased blood lipid levels.

Due to the significant impact of a memory-related outcome such as dementia on exposure recall, case-control studies were excluded, though nested case-control studies, where historical records are used to determine the exposure status, were eligible for inclusion. Cross-sectional studies, qualitative studies, case reports/series and narrative reviews were also excluded, as were studies that measure change in continuous cognitive measures (e.g. MoCA score) without attempt to map these scores to ordinal groups (e.g. no dementia/dementia). Previous systematic reviews were not eligible for inclusion, but their reference lists were screened to identify any potentially relevant articles.

@@

How to interpret agreement co-efficients is widely debated, and while arbitary cut-off values may mislead readers,[@brennan1992] they provide a useful rubric by which to assess inter-rater agreement. Here, I used guidelines based on a stricter interpretation of the Cohen's $kappa$ coefficient,[@mchugh2012] presented in Table \@ref(tab:gwet-table).

@@

@@

\begin{equation}
Mean = \frac{(N_1M_1 + N_2N_2)}{(N_1 + N_2)}
  (\#eq:combiningGroups2)
\end{equation}

@@

<!-- TODO This might not be as accurate now given the focus on dose response --> <!-- TODO Cross reference with the data extraction issues here - i.e. not providing qratile cut-off values, not poviding numbers per group, etc. --> Where necessary, lipid levels reported in _mmol/L_ were converted in _mg/dL_ using the following formula:

@@

where $Z = 38.67$ for total cholesterol, LDL-c and HDL-c, and $Z = 88.57$ for triglycerides. For widely-used categorises of lipids levels on the _mg/dL_ scale, see Table \@ref(tab:lipidLevels-table) in Section \@ref(intro-lipid-fractions).

@@

The range of effect measures presented by studies (odds ratios, risk ratios, hazard ratios, etc) are not directly interchangeable in the context of systematic review.  As such, different effect estimates can be one potential problem that precludes a meta-analysis of all studies.[@mckenzie2019] If the outcome is rare, as is the case for dementia outcomes, the estimated prevalence of  odds and risk ratios will approximate each other.<!-- TODO CITATION NEEDED --> However, hazard ratios provide a very different interpretation, taking into account person-time-at-risk in each treatment group.

@@

This move was prompted by a unclear definition of "methodological quality" which could include facets such as unclear reporting, in addition to challenges in the comparison of results from different tools. <!-- TODO CITATION NEEDED --> As part of this shift, the focus shifted from checklist or score based tools towards domain-based methods, in which different potential sources of bias in a study are assessed in order. <!-- TODO CITATION NEEDED -->
Finally, the new tools tools move from assessing bias at the study level to considering separately each individual numerical result reported.  For example, a study may report on the efficacy of an intervention at six months and two years follow-up. In this case, missing outcome data that is not an issue at six months may introduce bias after two years of follow-up, and assigning a single risk-of-bias judgement to the study as a whole masks the different biases applicable to each unique result.

@@

Randomized controlled trials were assessed using the RoB 2 tool.[@sterne2019] The tool assess the risk of bias across five domains: bias arising from the randomization process, bias due to deviations from intended intervention, bias due to missing outcome data, bias in measurement of the outcome, bias in selection of the reported result. Acceptable judgements for each domain include: "low risk", "some concerns", "high risk". Each of the five domains contains a series of signalling questions or prompts which guide the user through the tool. Once a domain-level judgement for each domain has been assigned, an overall judgement, using the same three levels of risk of bias, is assigned to the result.

@@

For non-randomised studies of interventions (NRSI), I used the ROBINS-I (Risk Of Bias In Non-randomised Studies - of Interventions) tool.[@sterne2016] This tool assess the risk of bias across seven domains: bias due to confounding, bias due to selection of participants, bias in classification of interventions, bias due to deviations from intended interventions, bias due to missing data, bias in measurement of outcomes, and bias in selection of the reported result. Similar to RoB 2, it has a number of prompting questions per domain, with acceptable judgements including “low risk”, “moderate risk”, “serious risk” and “critical risk”. In the context of the tool, observational studies are assessed in reference to an idealised randomised controlled trial. Under this approach, the (rare) overall judgement of "Low" indicates that the results should be considered equivalent to produced by a randomised controlled trial.

While a risk-of-bias tool for non-randomised studies of exposures (NRSE) is currently under development,[@morganr2020] but was insufficiently developed at the time the risk-of-bias assessments for this review were performed. Instead, I used a version of the ROBINS-I tool informed by the preliminary ROBINS-E tool (Risk of Bias In Non-randomised Studies – of Exposure), which I had previously applied in a published review.[@french2019] The version had no signalling questions and so judgements, using the same four levels of bias as ROBINS-I, were made at the domain level. The motivation using this tool above other established tools such as the Newcastle-Ottowa scale (NOS)[@wells2000] was two-fold. In the first instance, as mentioned in the introduction to this section, using a domain-based tool has distinct advantages over better-developed checklist-type tools including the NOS. Additionally, using a domain-based tool for non-randomised studies of exposures enabled better comparison with risk-of-bias assessments performed for the other study designs as part of this review.

@@

At present, no formalised risk-of-bias assessment tool for Mendelian randomization studies is available. Assessment of the risk of bias in Mendelian randomisation studies was informed by the approach used in a previous systematic review of Mendelian randomisation,[@mamluk2020] as identified by a review of risk-of-bias assessments in systematic reviews of Mendelian randomisation studies (advance results from this review were obtained from contact with the review authors).[@spiga2021] A copy of this tool is available in Appendix \@ref(appendix-mr-rob), but in summary, results were assessed for bias arising from weak instruments, genetic and other confounding, pleiotropy and population stratification. Acceptable judgements for each fo the 5 domains in the tool included "low", "moderate" and "high" risk of bias.

@@

In addition to assessing the risk of bias within each result contributing to a synthesis, I also assessed risk of bias due to missing evidence at the analysis level. This assessment examines evidence missing due to selective non-reporting - as distinct from the selective reporting of a single result from multiple planned - and was performed using the forthcoming RoB-ME (Risk of Bias due to Missing Evidence in a synthesis) tool.[@zotero-15123] <!-- TODO Need more here describing missing evidence --> The tool is in development stages, and as part of this review, I piloted the tool, and provided feedback to the developers. <!-- TODO Cite Ben's work on effect of missing study versus choice of analysis method -->

@@

An inital qualiatative synthesis of evidence was performed, summarising the data extracted from studies stratified by study design. Where individual studies were deemed comparable, they were incorporated into a quantitative analysis or "meta-analysis". <!-- TODO CITATION NEEDED -->

@@

Results were stratified into subgroups on the basis of the overall risk of bias assessment, and summary estimates for each subgroup, in addition to an overall effect estimate, are displayed in each forest plot. Additional descriptive statistics are presented, while prediction intervals are shown as a dotted line banding the overall effect estimate. Finally, where at least 10 results are available, a test subgroup differences between studies at different levels of risk of bias was performed (see subsequent Section \@ref(sys-rev-visualising-results)). All models were implemented using the `metafor` R package.

@@

Several of the included studies presented data on multiple categories of lipid levels, but provided an overall effect estimate based on a comparison of only two of these categories (e.g. for example, highest vs lowest quartile). <!-- TODO Question: if I look at highest vs lowest, and also at top dose vs lowest dose, am I double counting results? --> While this allows for easy interpretation of the resulting effect estimate, it ignores any potential non-linear relationships between the exposure and outcome, in addition to discarding useful information contain in the interim groups. In order to address this limitation, I performed a dose-response meta-analysis in those studies reporting more than two categories for lipid levels. This marks a departure from the published protocol, as I was unaware that such a large number of studies would report dementia risk across multiple lipid categories.

Studies were excluded from this analysis if the number of categories was less than three or if the necessary information for synthesis (cut-off points, number of participants and number of events per category) was not available. A restricted cubic spline model was fitted to allow for a non-linear relationship, for example a U or J-shaped relationship, where low and high levels of the exposure can have different effects versus a "normal" reference dose. <!-- TODO CITATION NEEDED --> The locations of the knots in the model wer identified using fixed percentiles (25th, 50th, 75th) of the exposure data. Reference doses were defined _a priori_ as the cut-off of the "Normal"/"Optimal" categories for each fraction, as detailed in Table \@ref(tab:lipidLevels-table). Under this approach, the reference dose was defined as 200 mg/dL for total cholesterol, 100 mg/dL for LDL-c, 40 mg/dL for HDL-c, and 150 mg/dL for triglycerides.

@@

Where there was evidence of heterogeneity between results included in a meta-analysis, I investigated this further using meta-regression against reported characteristics. _A priori_, I was interested in the effect that the age at baseline, sex and risk-of-bias judgement had on the results. Syntheses with greater than 10 studies were assessed for heterogeneity across these covariates. <!-- TODO Need more here -->

Finally, I investigated the potential for small study effects, which may be caused by publication bias, both visually using funnel plots and formally using Egger's regression test. <!-- TODO CITATION NEEDED - cite Jonathon on interpreting small study effects, and whatever citation is needed for  -->

@@

One of the limitations of current risk-of-bias assessments in systematic reviews is that they are often divorced from the results to which they refer, and are infrequently incorporated into the analysis. <!-- TODO CITATION NEEDED --> In response to this criticism, I developed a new visualisation tool was designed to allow for the production of "paired" forest plots, as recommended by the ROB2 publication, where a risk-of-bias assessment is presented alongside it's corresponding numerical results.[@sterne2019] This tool was developed as an adjunct to this thesis to aid in creating standardised risk-of-bias figures,[@mcguinness2020robvisPaper] and the "paired" forest plot functionality grew out of a collaboration with other researchers to design a modular method for creating custom forest plots.[@zotero-14999] <!-- Apparently collaboration is a big thing they want to see --> A summary of this tool is contained in Appendix \@ref(appendix-robvis), and all forest plots presented in this Chapter were created using this tool.

@@

_[Note: Julian, I am particularly interested in your feedback on this section, and the corresponding results section (Section \@ref(sys-rev-including-preprints-res)), as I am not convinced on the language I am using]_

@@

@@

## Results

<!-- TODO Compare findings for two papers from Rantanen - not sure why? -->

### Initial search and validation of search filters

The database search identified `r num_to_text(23447)` records, of which `r comma(7338)` were duplicated records. Of the random sample of 500 records screened to ensure the accuracy of the study design filters, no eligible records were identified. Many of those excluded by the filters were commentaries/educational articles, or described basic science studies. <!-- TODO include some example citations here -->

### Screening results

Following de-duplication, the titles and abstracts of `r num_to_text(16109)` records were assessed for eligibility. `r num_to_text(387)` were deemed potentially eligible, and the full text records for these were accessed and screened.

The PRISMA flow diagram <!-- TODO CITATION NEEDED - cite PRIMSA flow paper here, if published --> presented in Figure \@ref(fig:prisma-flow-fig), illustrates the movement of articles through the review. To highlight the contribution of preprint archives to the review, the flow diagram delineates between those records captured through databases searches (presented on the left of the diagram) and those captured by the search tool described in the previous chapter (presented in grey on the right of the diagram).

Common reasons for exclusion at the full text-stage included studies that reported on the wrong exposures (n = `r n_wrong_exposure` ; most commonly a ineligible lipid fraction), used the wrong study design (n= `r n_study_design`), or reported on a wrong outcome (n=24; e.g. change in cognitive scores).

<!----------------------------------------------------------------------------->

\blandscape{}

(ref:prisma-flow-cap) PRISMA flow diagram illustrating how records moved through the systematic review process. The different contributions of standard bibliographic databases and preprint servers to the review are indicated.

(ref:prisma-flow-scap) PRISMA flow diagram

```{r prisma-flow-fig, echo = FALSE, results="asis", fig.pos="H", fig.cap='(ref:prisma-flow-cap)', out.width='100%', fig.scap='(ref:prisma-flow-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/prismaflow.png"))
```

\elandscape{}

<!----------------------------------------------------------------------------->

### Validation of screening

```{r agree-setup, include = FALSE}
```

For the assessment of the intra-/inter-rater reliability, the estimated values of $AC1$ were interpreted against the categories presented in Table \@ref(tab:gwet-table). For the inter-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeInter_coeff[1]`, $kappa$ = `r agreeInter_coeff[2]`, Table \@ref(tab:agreeInter-table)). Similarly for intra-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeIntra_coeff[1]`, $kappa$ = `r agreeIntra_coeff[2]`, Table \@ref(tab:agreeIntra-table)). The discrepancy between the $AC1$ and $kappa$ coefficients illustrates the sensitivity of $kappa$ to imbalanced marginals, caused in this sample by a large imbalance  towards exclusion.[@feinstein1990]

<!----------------------------------------------------------------------------->
(ref:agreeInter-caption) Inter-rater agreement on a subset of records, indicating high accuracy.

<!-- TODO Include AC1 result in caption above -->

(ref:agreeInter-scaption) Inter-rater agreement

```{r agreeInter-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

<!----------------------------------------------------------------------------->
(ref:agreeIntra-caption) Intra-rater agreement on subset of records, indicating high consistency.

(ref:agreeIntra-scaption) Inter-rater agreement

```{r agreeIntra-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

Those records which were excluded in the initial screening, but were included by the second reviewer (n=`r discrepancy_Inter`, Table \@ref(tab:agreeInter-table)) were investigated. This discrepancy between the two reviewers was explained in all cases by differing interpretations of the inclusion criteria, most commonly around the definition of cognitive decline versus dementia and the definition of eligible lipids fractions.

### Characteristics of included studies {#sys-rev-characteristics}

```{r characteristicsSetup, message=FALSE, results="asis", echo = FALSE}
```

Following full-text screening, `r n_included` unique studies (described across `r n_reports_included` reports) met the criteria for inclusion in the review.`r get_all_citations()` Table \@ref(tab:studyCharacteristics-table) presents a summary of the characteristics of each study.

The majority of included studies described non-randomised analyses, with the sole two included randomised controlled trials (the Heart Protection Study/British Heart Foundation trial[@heartprotectionstudycollaborativegroup2002] and the JUPITER trial[@ridker2008]) both reporting on the effect of statin use on all-cause dementia in older adults. A similarly small number of Mendelian randomisation studies were identified, several of which employed a two-sample approach using summary statistics from the same published genome wide association studies (GWAS) leading to complications in the synthesis (see Section \@ref(sys-rev-res-AD)).

Of the `r n_nrsi_included` non-randomised studies examining treatments that modify lipid levels, all examined statin use (`r n_statins`), while a small number also reported on other non-statins agents such as fibrates (`r n_fibrates`). In the `r n_nrse_included` non-randomised studies of exposure, hypercholesterolemia (`r n_hyperchol`) and total cholesterol levels (`r n_TC`) were the most frequently reported risk factors.

In terms of outcomes, the vast majority of studies examined either all-cause dementia (`r n_Dementia`) or Alzheimer's disease (`r n_AD`), with only a small proportion examining vascular dementia (`r n_VaD`). Some other outcome classifications such as vascular-component or mixed dementia were also investigated, but were much rarer.

<!-- Many studies using electronic health records as their data source did not accurately report the diagnostic codes used to identify cases. -->

<!-- Midlife cholesterol levels were of particular interest, with several studies examining this age-range.  -->

<!--- TODO Talk a little bit more about age here --->

Three included reports were preprinted (denoted in the Table \@ref(tab:studyCharacteristics-table) using an asterisk),[@ridker2008; @so2017;@zhu2017] one of which had subsequently been published and was captured by the primary literature search.[@zhu2018] All three included preprints were obtained from the bioRxiv preprint server and all described a Mendelian randomisation analysis.

<!-- As discussed in Section \@ref(rev-discussion-MR), the choice was not material, as the results from all studies using these data sources were comparable. -->

<!-- TODO Figures and tables to include here: -  Summary of risk of bias (not done) -->

<!-- TODO Also need to extract the codes used in each study -->

<!-- TODO Need to add abbreviations to table -->

<!-- TODO Ensure citations make sense, and are correct! Get first reference and then add -->

\blandscape{}
<!----------------------------------------------------------------------------->
(ref:studyCharacteristics-caption) Characteristics of included studies, stratified by study design. Note that three studies reported on multiple analytical designs within a single study report, and these have been duplicated across the relevant sub-sections.

(ref:studyCharacteristics-scaption) Characteristics of included studies

(ref:studyCharacteristics-first-cite) [@heartprotectionstudycollaborativegroup2002]

```{r studyCharacteristics-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->
\elandscape{}

```{r cohortLocationsSetup, include = FALSE}
```

As illustrated in Figure \@ref(fig:cohortLocations), the majority of reports described studies conducted in high-income countries, with the most high represented region being North America. Of interest, several of the included studies were conducted in Taiwan (`r n_taiwan`), all but one of which made use of the Taiwan National Health Insurance database.

<!----------------------------------------------------------------------------->

(ref:cohortLocations-cap) Geographical distribution of study cohorts

(ref:cohortLocations-scap) Geographical distribution of study cohorts

```{r cohortLocations, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:cohortLocations-cap)', out.width='100%', fig.scap='(ref:cohortLocations-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/cohortLocations.png"))
```
<!----------------------------------------------------------------------------->
&nbsp;

Finally, there were several eligible studies reported as conference abstracts that did not present numerical results. These reports were included in the analysis to enable assessment of risk of bias due to missing evidence (see Section \@ref(methods-rob-me)).

### Risk of bias {#risk-of-bias-res}

```{r riskOfBias, include = FALSE, message=FALSE, warning=FALSE}
```

As discussed above, the risk-of-bias assessments are presented alongside their corresponding numerical result. A more detailed discussion of the sources and directions of bias is presented in Chapter \@ref(tri-heading), and so this section presents a brief summary of the biases observed in each study design.

For the two randomised controlled trials, both were judged to be at low risk of bias. In contrast, many of the non-randomised studies of statin use were at serious risk of bias due poor controlling for confounding, immortal time bias, and missing outcome data (as noted above, a judgement of low risk of bias for a non-randomised study is rare). Similarly, non-randomised studies of exposures suffered from incomplete adjustment for potentially important confounders, and concerns over the selection of the reported result from among several analyses (e.g. examination of lipids as a binary or continuous variable). Finally, bias was introduced into Mendelian randomisation studies via the potential for hoziontal pleiotropy and population stratification.[@davies2018]

Following best practice, any result judged to be at critical risk of bias should be excluded from any quantitative analyses. <!-- TODO CITATION NEEDED --> `r num_to_text(n_critical,T)` observational studies were excluded on this basis, predominantly due to a lack of adjustment for any potentially important confounders (i.e. the study reported unadjusted estimates).`r critical_citations`

### All-cause dementia {#sys-rev-res-Dementia}

#### Statins

```{r rctStatinDementia, include = FALSE, message=FALSE, warning=FALSE}
```

>>>>

# Systematic review of existing evidence on the association between blood lipids and dementia outcomes: Methods {#sys-rev-methods-heading}

@@

@@

This chapter presents the methods used to perform a systematic review of primary studies that have examined the relationship between the levels of blood lipids (such as cholesterol and triglycerides) and dementia outcomes. In addition, the review examined the relationship betweeen treatments that change blood lipid levels, such as statins, and dementia outcomes. The results of this systematic review are then presented in Chapter \@ref(sys-rev-results-heading.)

@@

In this chapter, I describe a comprehensive systematic review of the relationship between blood lipid levels (and treatments that modify them) and the subsequent risk of dementia and related outcomes.

This analysis sought to address two specific aims. Firstly, as discussed in the Introduction to this thesis (Section \@ref(evidence-association)), several diverse forms of evidence on the relationship of lipids and dementia exist. These include randomised controlled trials, observational studies of different design, and Mendelian randomisation studies. However, based on a scoping review of existing literature, no previous evidence synthesis exercise has attempted to examine the association of lipids/statins with dementia outcomes across these distinct evidence types. Collating these diverse evidence sources is important, as if the observed association between lipids and dementia is constant across them, it increases our confidence in the association. As such, the primary aim of this analysis was to systematically review all available literature describing prospective analyses, regardless of study design.

@@

Given the size of the review, I have seperated the methods and results into two chapters for ease of reading. This chapter details the methodology used, while CHapter \@ref(sys-rev-results-heading) presents the results of the review.

@@

A pre-specified protocol for this analysis was registered on the Open Science Framework platform and is available for inspection.[@mcguinnessluke2020]

In line with best-practice guidance, secondary reviewers were used to check the accuracy of screening, data extraction and risk-of-bias assessment processes. Due to the scale of the project, this review was performed in conjunction with a team of secondary reviewers (see Acknowledgements and Author Declaration in the front matter).

@@

### Eligibility criteria

I sought to include studies that examined blood lipid levels as a risk factor for dementia outcomes, defined either as binary hypercholesterolemia variable or by category/1-standard-deviation increase of a specific lipid fraction (total cholesterol, high- and low-density lipoprotein cholesterol and triglycerides). I also aimed to include studies examining the effect of treatments that modify lipids levels as a source of indirect evidence. Eligible study designs included randomized controlled trials and non-randomized observational studies of lipid modifying treatments, longitudinal studies examining the effect of increased/decreased blood lipid levels, and genetic instrumental variable (Mendelian randomization) studies examining the effect of genetically increased/decreased blood lipid levels.

<!-- COMMENT Isn’t more true to say it was assumed that a health care professional had made the diagnosis (unlikely to use specific criteria) -->

@@

Due to the significant impact of a memory-related outcome such as dementia on exposure recall, case-control studies were excluded, though case-control studies where historical records are used to determine the exposure status, were eligible for inclusion. Cross-sectional studies, qualitative studies, case reports/series and narrative reviews were also excluded, as were studies that measure change in continuous cognitive measures (e.g. MoCA score) without an attempt to map these scores to ordinal groups (e.g. no dementia/dementia). Previous systematic reviews were not eligible for inclusion, but their reference lists were screened to identify any potentially relevant articles.

### Information sources and search strategy

@@

Grey literature was searched via ProQuest, OpenGrey and Web of Science Conference Proceedings Citation Index, while theses were accessed using the Open Access Theses and Dissertations portal. In addition, the abstracts list of relevant conferences (e.g. the proceedings of the Alzheimer's Association International Conference, published in the journal Alzheimer's & Dementia) were searched by hand. <!-- TODO how were these searched? Date and other range --> Finally, the reference lists of included studies were searched by hand while studies citing included studies was examined using Google Scholar (forward and reverse citation searching or "snowballing"[@greenhalgh2005; @wohlin2014]).

@@

The criteria against which records were assessed for eligibility are presented above.

@@

How to interpret agreement coefficients is widely debated, and while arbitrary cut-off values may mislead readers,[@brennan1992] they provide a useful rubric by which to assess inter-rater agreement. Here, I used guidelines based on a stricter interpretation of the Cohen's $kappa$ coefficient,[@mchugh2012] presented in Table \@ref(tab:gwet-table).

@@

<!-- COMMENT [Yoav] How did you choose what data to include? re: studyification -->

@@

\begin{equation}
Mean = \frac{(N_1M_1 + N_2M_2)}{(N_1 + N_2)}
  (\#eq:combiningGroups2)
\end{equation}

@@

Where necessary, lipid levels reported in _mmol/L_ were converted in _mg/dL_ using the following formula:

@@

where $Z = 38.67$ for total cholesterol, LDL-c and HDL-c, and $Z = 88.57$ for triglycerides. The choice of _mg/dL_ was influenced by the widely-used categorises of lipids levels on the _mg/dL_ scale, as shown in Table \@ref(tab:lipidLevels-table) in Section \@ref(intro-lipid-fractions).

@@

The range of effect measures presented by studies (odds ratios, risk ratios, hazard ratios, etc) are not directly interchangeable in the context of systematic review.  As such, different effect estimates can be one potential problem that precludes a meta-analysis of all studies.[@mckenzie2019] If the outcome is rare, as is the case for dementia outcomes, the estimated odds and risk ratios will approximate each other.<!-- TODO CITATION NEEDED --> However, hazard ratios provide a very different interpretation, taking into account person-time-at-risk in each treatment group.

@@

This move was prompted by an unclear definition of "methodological quality" which could include facets such as unclear reporting, in addition to challenges in the comparison of results from different tools. <!-- TODO CITATION NEEDED --> As part of this shift, the focus shifted from checklist or score based tools towards domain-based methods, in which different potential sources of bias in a study are assessed in order. <!-- TODO CITATION NEEDED -->
Finally, the new tools tools move from assessing bias at the study level to considering separately each individual numerical result reported.  For example, a study may report on the efficacy of an intervention at six months and two years follow-up. In this case, missing outcome data that is not an issue at six months may introduce bias after two years of follow-up, and assigning a single risk-of-bias judgement to the study as a whole masks the different biases applicable to each unique result.

@@

Randomized controlled trials were assessed using the RoB 2 tool.[@sterne2019] The tool assesses the risk of bias across five domains: bias arising from the randomization process, bias due to deviations from intended intervention, bias due to missing outcome data, bias in measurement of the outcome, bias in selection of the reported result. Acceptable judgements for each domain include: "low risk", "some concerns", "high risk". Each of the five domains contains a series of signaling questions or prompts which guide the user through the tool. Once a domain-level judgement for each domain has been assigned, an overall judgement, using the same three levels of risk of bias, is assigned to the result.

@@

For non-randomised studies of interventions (NRSI), I used the ROBINS-I (Risk Of Bias In Non-randomised Studies - of Interventions) tool.[@sterne2016] This tool assess the risk of bias across seven domains: bias due to confounding, bias due to selection of participants, bias in classification of interventions, bias due to deviations from intended interventions, bias due to missing data, bias in measurement of outcomes, and bias in selection of the reported result. Similar to RoB 2, it has a number of prompting questions per domain, with acceptable judgements including “low risk”, “moderate risk”, “serious risk” and “critical risk”. In the context of the tool, observational studies are assessed in reference to an idealised randomised controlled trial. Under this approach, the (rare) overall judgement of "Low" indicates that the results should be considered equivalent to that produced by a randomised controlled trial.

While a risk-of-bias tool for non-randomised studies of exposures (NRSE) is currently under development,[@morganr2020] but was insufficiently developed at the time the risk-of-bias assessments for this review were performed. Instead, I used a version of the ROBINS-I tool informed by the preliminary ROBINS-E tool (Risk of Bias In Non-randomised Studies – of Exposure), which I had previously applied in a published review.[@french2019] The version had no signaling questions and so judgements, using the same four levels of bias as ROBINS-I, were made at the domain level. The motivation using this tool above other established tools such as the Newcastle-Ottowa scale (NOS)[@wells2000] was two-fold. In the first instance, as mentioned in the introduction to this section, using a domain-based tool has distinct advantages over better-developed checklist-type tools including the NOS. Additionally, using a domain-based tool for non-randomised studies of exposures enabled better comparison with risk-of-bias assessments performed for the other study designs as part of this review.

@@

At present, no formalised risk-of-bias assessment tool for Mendelian randomization studies is available. Assessment of the risk of bias in Mendelian randomisation studies was informed by the approach used in a previous systematic review of Mendelian randomisation,[@mamluk2020] as identified by a review of risk-of-bias assessments in systematic reviews of Mendelian randomisation studies (advance results from this review were obtained from contact with the review authors).[@spiga2021] A copy of this tool is available in Appendix \@ref(appendix-mr-rob), but in summary, results were assessed for bias arising from weak instruments, genetic and other confounding, pleiotropy and population stratification. Acceptable judgements for each of the 5 domains in the tool included "low", "moderate" and "high" risk of bias.

@@

In addition to assessing the risk of bias within each result contributing to a synthesis, I also assessed risk of bias due to missing evidence at the analysis level. This assessment examines evidence missing due to selective non-reporting - as distinct from the selective reporting of a single result from multiple planned analyses - and was performed using the forthcoming RoB-ME (Risk of Bias due to Missing Evidence in a synthesis) tool.[@zotero-15123] <!-- TODO Need more here describing missing evidence --> The tool is in development stages, and as part of this review, I piloted the tool, and provided feedback to the developers.

@@

An initial qualitative synthesis of evidence was performed, summarising the data extracted from studies stratified by study design. Where individual studies were deemed comparable, they were incorporated into a quantitative analysis or "meta-analysis". <!-- TODO CITATION NEEDED -->

@@

Results were stratified into subgroups on the basis of the overall risk of bias assessment, and summary estimates for each subgroup, in addition to an overall effect estimate, are displayed in each forest plot. Additional descriptive statistics are presented, while prediction intervals are shown as a dotted line banding the overall effect estimate. Finally, where at least 10 results are available, a test of subgroup differences between studies at different levels of risk of bias was performed (see subsequent Section \@ref(sys-rev-visualising-results)).[@deeks2019] All models were implemented using the `metafor` R package.

@@

Several of the included studies presented data on multiple categories of lipid levels, but provided an overall effect estimate based on a comparison of only two of these categories (e.g. for example, highest vs lowest quartile). <!-- TODO Question: if I look at highest vs lowest, and also at top dose vs lowest dose, am I double counting results? --> While this allows for easy interpretation of the resulting effect estimate, it ignores any potential non-linear relationships between the exposure and outcome, in addition to discarding useful information contained in the interim groups. In order to address this limitation, I performed a dose-response meta-analysis in those studies reporting more than two categories for lipid levels. This marks a departure from the published protocol, as I was unaware that such a large number of studies would report dementia risk across multiple lipid categories.

Studies were excluded from this analysis if the number of categories was less than three or if the necessary information for synthesis (cut-off points, number of participants and number of events per category) was not available. A restricted cubic spline model was fitted to allow for a non-linear relationship, for example a U or J-shaped relationship, where low and high levels of the exposure can have different effects versus a "normal" reference dose. <!-- TODO CITATION NEEDED --> The locations of the knots in the model were identified using fixed percentiles (25th, 50th, 75th) of the exposure data. Reference doses were defined _a priori_ as the cut-off of the "Normal"/"Optimal" categories for each fraction, as detailed in Table \@ref(tab:lipidLevels-table). Under this approach, the reference dose was defined as 200 mg/dL for total cholesterol, 100 mg/dL for LDL-c, 40 mg/dL for HDL-c, and 150 mg/dL for triglycerides.

@@

Where there was evidence of heterogeneity between results included in a meta-analysis, I investigated this further using meta-regression against reported characteristics. _A priori_, I was interested in the effect that the age at baseline, sex and risk-of-bias judgement had on the results. Syntheses with greater than 10 results were assessed for heterogeneity across these covariates.[@deeks2019]

Finally, I investigated the potential for small study effects, which may be caused by publication bias, both visually using funnel plots and formally using Egger's regression test.[@sterne2011]

@@

One of the limitations of current risk-of-bias assessments in systematic reviews is that they are often divorced from the results to which they refer, and are infrequently incorporated into the analysis.[@marusic2020; @katikireddi2015] In response to this criticism, I developed a new visualisation tool which was designed to allow for the production of "paired" forest plots - a risk-of-bias assessment is presented alongside it's corresponding numerical result -  as recommended by the ROB2 publication.[@sterne2019] This tool was developed as an adjunct to this thesis to aid in creating standardised risk-of-bias figures,[@mcguinness2020robvisPaper] and the "paired" forest plot functionality grew out of a collaboration with other researchers to design a modular method for creating custom forest plots.[@zotero-14999] A summary of this tool is contained in Appendix \@ref(appendix-robvis), and all forest plots presented in this Chapter were created using this tool.

@@

<!-- TODO _[Note: Julian, I am particularly interested in your feedback on this section, and the corresponding results section (Section \@ref(sys-rev-including-preprints-res)), as I am not convinced on the language I am using]_ -->

@@

<!-- COMMENT You need to state a priori in what objective way you can judge if pre-prints do or do not have added value -->

@@

## Summary

* In this chapter, I have presented the methods underpinning a comprehensive systematic review of the existing literature on the association of dementia incidence with blood lipids and treatments for lowering blood lipids.

* I have highlighted how this review will make use of preprinted literature, using the research tool introduced in the preceding chapter (Chapter \ref(sys-rev-tools-heading)).

* In the following chapter, I presented the results of this comprehensive systematic review and detail how the identified evidence is used throughout the remainder of the thesis.
